---
title: "Multivariate General Linear Models"
subtitle: "Lecture 08.1: Multivariate Overview"
author:  "Lauren Sullivan"
date: "Module: Multivariate Models"
output: beamer_presentation
mainfont: Calibri 
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

rm(list=ls())
library(tidyverse)
library(scatterplot3d)  # for 3D point figure.
library(GGally)  #for scatmat plot and correlation plots
library(PerformanceAnalytics)  #for creating the correlation plots with values on them


sparrow <- read_csv("../data/sparrow.csv")

```



## Readings

\textbf{Required for class:}

  - NA

\bigskip\textbf{Optional:}

- [\textcolor{teal}{Healy. (1969) 259 Note: Rao's Paradox Concerning Multivariate Tests of Significance. \textit{Biometrics}.}](https://www.jstor.org/stable/pdf/2528798.pdf?casa_token=UZEX4Zq4I4cAAAAA:UbwfoIiNum8snVaBC6XCvKs7x5qfhVLiu2j7oBhtpf0Z8Zg4bgOagMLlz60QK2iNyGuP2M9xog0MXW28utbHzd2gB29zTEgv8jcl7ZAJ5ttsjqWh)



## Univariate Analysis

When analyzing univariate data (what we have been doing up till now), we have been learning how to explain the variation in a single dependent variable (Y).

\scriptsize
```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=6}

ggplot(sparrow)+
  geom_histogram(aes(x=AE), binwidth = 1)+  
  theme_bw()+
  theme(axis.title = element_text(face="bold", size=16))


```


## Multivariate Analysis

When analyzing multivariate data, we will attempt to explain variation in more than one dependent variable (multiple Y's).

\bigskip

\scriptsize
```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=6}
scatterplot3d(sparrow[5:7])
```

## Why Multivariate?

Multivariate analysis allows you to look at a more complex description of patterns in biology. In many cases, separate univariate analyses miss a covariation signal in the data, so it's important to think about data more fully if possible.

\bigskip

There are several ways to look at multivariate patterns from a matrix of **Y**'s.

1.  Linear models: MANOVA/regression to test patterns
2.  Ordination: PCA, nMDS, etc to visualize patterns
3.  Permutation tests: PERMANOVA to test patterns


## Rao's Paradox and Data Collection

Increasing the number of dimensions of your data (# of Y's) means more information about a process, but for a given number of observations (rows of your dataset) the statistical power decreases.  You need to watch out to make sure you don't have too few *n*'s for the number of dependent variables (Y's).

\bigskip
There are a lot of suggestions for the number of observations you need.  Make sure you think about this consciously!

- $n = 2*(\# Y)$
- $n = 4*(\# Y)$
- $n = (\# Y)^2$
- $n_{group} = 2*(\# Y)$
- $n_{group} = 4*(\# Y)$



## Multivariate Data

Multivariate data is can be described as:

\begin{center}
$\bf{Y} \sim MVN(\bf{\mu},\Sigma)$  
\end{center}

\bigskip
With the Variance-Covariance (VCV) matrix that is described as:

$$
\bf{\Sigma} = \begin{bmatrix}
s_{11} &  &  \\
s_{21} & s_{22} &  \\
s_{31} & s_{32} & s_{33}\\
\end{bmatrix}
$$


## Correlation Matrix


The VCV can be standardized to be the correlation matrix (last time was $\rho$ but I've replaced with 
$r$'s). First you pull the $\sigma_i^2$ out of the matrix.
$$
\bf{\Sigma} = \sigma_i^2\begin{bmatrix}
1 &  & \\
 r_{21} & 1 &  \\
 r_{31} & r_{32} & 1 \\
\end{bmatrix}
$$

\bigskip
Then you can simply look at the correlation matrix (get rid of $\sigma_i^2$, and $\bf{\Sigma}$ simplifies to $\bf{R}$). 

$$
\bf{R} = \begin{bmatrix}
1 &  & \\
 r_{21} & 1 &  \\
 r_{31} & r_{32} & 1 \\
\end{bmatrix}
$$

## Plotting the Correlation Matrix

The correlation matrix can be visualized by plotting the correlation between variables.

\bigskip
\scriptsize
```{r, eval=TRUE, echo=TRUE, warning=FALSE, message=FALSE, fig.height=5}

chart.Correlation(sparrow[,5:7], histogram = TRUE, pch = 19)

```


## Another option in tidyverse

The `GGally` library and `ggscatmat()` also makes these figures

\bigskip
\scriptsize
```{r, eval=TRUE, echo=TRUE, warning=FALSE, message=FALSE, fig.height=5}

ggscatmat(sparrow, columns = 5:7, alpha = 1)+
  theme_bw()+
  theme(axis.title = element_text(face="bold", size=16))

```










