---
title: "Data Transformation and Exploration"
subtitle: "Lecture 3.4: Data Simulation"
author:  "Lauren Sullivan"
date: "Module: Data Management, Visualization & Reproducibility"
output: beamer_presentation
mainfont: Calibri 
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(cowplot)
theme_set(theme_cowplot())

```



## Data Simulation

Sometimes you need to simulate data to determine:

- if your models are working like you think they are
- to determine appropriate sample size (e.g. power analysis)
- to understand your data better


## the 4 types of simulator functions in R

R has 4 built in functions for most distributions that you will find useful when simulating data.

- `r` - for generating random numbers from a distribution
- `d` - for density which evaluates a distribution at a certain point
- `p` - for evaulating the cumulative distribution function for a given distribution
- `q` - for the quantile function

## `rnorm()`

```{r,  echo=TRUE, warning=FALSE, message=FALSE}
rnorm(10, mean = 1, sd = 0.5)

```


## `rnorm()`

```{r,  echo=TRUE, warning=FALSE, message=FALSE}
rnorm(10, mean = 1, sd = 0.5)
rnorm(10, mean = 1, sd = 0.5)

```



## `set.seed()`

```{r,  echo=TRUE, warning=FALSE, message=FALSE}
set.seed(10)
rnorm(10, mean = 1, sd = 0.5)
set.seed(10)
rnorm(10, mean = 1, sd = 0.5)

```


## Plot the distribution
\scriptsize
```{r,  echo=TRUE, warning=FALSE, message=FALSE}
one <- rnorm(1000000, mean = 1, sd = 0.5)
dat <- tibble(one)

ten <- rnorm(1000000, mean = 5, sd = 2)
dat2 <- tibble(ten)
```

## Plot the distribution with `rnorm()`
\scriptsize
```{r,  echo=FALSE, warning=FALSE, message=FALSE}
ggplot(dat)+
  geom_density(aes(x = one), lwd = 2, lty = 1)+
  geom_density(aes(x = ten), lwd = 2, lty = 2)

```



## `pnorm()`

Calculate the probability of a certain point on a distribution

\scriptsize
```{r,  echo=TRUE, warning=FALSE, message=FALSE, fig.height=3.5, fig.width=6}

pnorm(2, mean = 1, sd = 0.5)
```

```{r,  echo=FALSE, warning=FALSE, message=FALSE, fig.height=3.5, fig.width=6}
data <- seq(-1, 3, length = 30)
data_t <- tibble(data)
data_t$dt <- dnorm(data, mean = 1, sd = 0.5)


dnorm_limit <- function(x) {
    y <- dnorm(x, mean = 1, sd = 0.5)
    y[x < -1  |  x > 2] <- NA
    return(y)
}


ggplot(data_t)+
  geom_line(aes(x = data, y = dt), lwd = 2, lty = 1)+
  stat_function(fun = dnorm_limit, geom = "area", fill = "blue", alpha = 0.2)+
  geom_vline(xintercept = 2,  color = "blue")






```

## `pnorm()`


\scriptsize
```{r,  echo=TRUE, warning=FALSE, message=FALSE, fig.height=3.5, fig.width=6}

pnorm(0.5, mean = 1, sd = 0.5)

```

```{r,  echo=FALSE, warning=FALSE, message=FALSE, fig.height=3.5, fig.width=6}

dnorm_limit <- function(x) {
    y <- dnorm(x, mean = 1, sd = 0.5)
    y[x < -1  |  x > 0.5] <- NA
    return(y)
}


ggplot(data_t)+
  geom_line(aes(x = data, y = dt), lwd = 2, lty = 1)+
  stat_function(fun = dnorm_limit, geom = "area", fill = "blue", alpha = 0.2)+
  geom_vline(xintercept = 0.5,  color = "blue")

```

## `dnorm()`

Find what the probability is at any point with the `dnorm()` function.

\scriptsize
```{r,  echo=TRUE, warning=FALSE, message=FALSE, fig.height=3.5, fig.width=6}

dnorm(1, mean = 1, sd = 0.5)

```
```{r,  echo=FALSE, warning=FALSE, message=FALSE, fig.height=3.5, fig.width=6}

ggplot(dat)+
  geom_density(aes(x = one), lwd = 2, lty = 1)+
  geom_hline(yintercept = dnorm(1, mean = 1, sd = 0.5),  color = "green")+
  geom_vline(xintercept = 1,  color = "blue")
```

## `dnorm()`

\scriptsize
```{r,  echo=TRUE, warning=FALSE, message=FALSE, fig.height=3.5, fig.width=6}

dnorm(2, mean = 1, sd = 0.5)

```
```{r,  echo=FALSE, warning=FALSE, message=FALSE, fig.height=3.5, fig.width=6}

ggplot(dat)+
  geom_density(aes(x = one), lwd = 2, lty = 1)+
  geom_hline(yintercept = dnorm(2, mean = 1, sd = 0.5),  color = "green")+
  geom_vline(xintercept = 2,  color = "blue")
```


## Plotting the distribution with `dnorm()`


```{r,  echo=FALSE, warning=FALSE, message=FALSE, fig.height=3.5, fig.width=6}

data <- seq(-1, 3, length = 30)
data_t <- tibble(data)
data_t$dt <- dnorm(data, mean = 1, sd = 0.5)
ggplot(data_t)+
  geom_line(aes(x = data, y = dt), lwd = 2, lty = 1)

```

## `qnorm()`

Used to find the probabilities of certain quantiles

\scriptsize
```{r,  echo=TRUE, warning=FALSE, message=FALSE, fig.height=3.5, fig.width=6}

qnorm(0.85, mean = 1, sd = 0.5)

```


```{r,  echo=FALSE, warning=FALSE, message=FALSE, fig.height=3.5, fig.width=6}


ggplot(data_t)+
  geom_line(aes(x = data, y = dt), lwd = 2, lty = 1)+
  geom_vline(xintercept = qnorm(0.85, mean = 1, sd = 0.5),  color = "blue")

```



## Other distributions
\scriptsize
```{r,  echo=TRUE, warning=FALSE, message=FALSE}
rpois(10, lambda = 1)

rbinom(10, size=12, prob=0.2)

rgamma(10, shape = 2, rate = 0.5)
```


## Simulate Data Based on Small Samples


\scriptsize

```{r,  echo=FALSE, warning=FALSE, message=FALSE}
dat <- tibble(ID = seq(1,20, 1), 
              mass = c(16.04601, 21.42249 ,19.06931, 20.54750 ,18.11079, 23.83877, 22.08287, 26.51201, 20.11507, 18.43206, 10.74246, 17.87718, 22.34310,  8.81798, 15.79852, 15.32697, 15.17099, 24.32051, 16.5276, 27.11078),
              sex = c(rep("F", 10), rep("M", 10))
                      )
```
```{r,  echo=TRUE, warning=FALSE, message=FALSE}
head(dat)

dat %>%
  group_by(sex) %>%
  summarize(mean_mass = mean(mass),
            sd_mass = sd(mass))



```


## Simulate Data Based on Small Samples


\scriptsize

```{r,  echo=TRUE, warning=FALSE, message=FALSE, fig.height=3.5, fig.width=6}

ggplot(dat)+
  geom_histogram(aes(x = mass, fill = sex))

```

## Simulate Data Based on Small Samples


\scriptsize

```{r,  echo=TRUE, warning=FALSE, message=FALSE, fig.height=3.5, fig.width=6}
set.seed(100)
dat_females <- rnorm(1000, mean = 20.61769, sd = 3.028194)
set.seed(24)
dat_males <- rnorm(1000, mean = 17.40361, sd = 5.748625)

dat_all <- tibble(c(dat_females, dat_males))
dat_all$sex <- c(rep("F", 1000), rep("M", 1000))
names(dat_all)[1] <- 'mass'

```

## Simulate Data Based on Small Samples

```{r,  echo=FALSE, warning=FALSE, message=FALSE, fig.height=3.5, fig.width=6}
ggplot(dat_all)+
   geom_density(aes(x = mass, color = sex), lwd = 2)


```


## Simulate a Linear Model

You may want to simulate from a specific model, instead of from a distribution of random numbers.  So say you want to simulate from this equation:


\smallskip
$y = \beta_0 + \beta_1x + \epsilon$, where $\epsilon \sim \mathcal{N}(0,2), \beta_0 = 0.5,$ and $\beta_1 = 2$


\scriptsize
```{r,  echo=TRUE, warning=FALSE, message=FALSE, fig.height=3.5, fig.width=6}
set.seed(20)

#simulate predictor (x) variables
x <- rnorm(100)

#simulate the error
e <- rnorm(100, 0, 2)

# calculate the response variables (y) from the model
y <- 0.5 + 2*x + e

summary(y)
```




## Simulate a Linear Model



\scriptsize
```{r,  echo=FALSE, warning=FALSE, message=FALSE, fig.height=3.5, fig.width=6}
dat <- as.tibble(cbind(x,y))
names(dat)[1] <- 'x'
names(dat)[2] <- 'y'


ggplot(dat)+
  geom_point(aes(x,y), size = 2)

```



## Simulate a Linear Model with a Binomial x

You could also use the same model, but instead have your x's distributed as a binomial variable (1's or 0's)



\scriptsize
```{r,  echo=TRUE, warning=FALSE, message=FALSE, fig.height=3.5, fig.width=6}
set.seed(20)

#simulate predictor (x) variables
x <- rbinom(100, 1, 0.5)
str(x)

#simulate the error
e <- rnorm(100, 0, 2)

# calculate the response variables (y) from the model
y <- 0.5 + 2*x + e

```



## Simulate a Linear Model with a Binomial x


\scriptsize
```{r,  echo=FALSE, warning=FALSE, message=FALSE, fig.height=3.5, fig.width=6}
dat <- as.tibble(cbind(x,y))
names(dat)[1] <- 'x'
names(dat)[2] <- 'y'


ggplot(dat)+
  geom_point(aes(x=x,y=y), size = 2)

```


## Simulate a Generalized Linear Model with Poisson

You may also want to simulate a model from something other than a normal distribution.  For example, say you want to use a Poisson log-linear model (often good for count data):


\smallskip
$Y \sim Poisson(\mu)$

$log(\mu) = \beta_0 + \beta_1x$, where $\beta_0 = 0.5,$ and $\beta_1 = 0.2$


\scriptsize
```{r,  echo=TRUE, warning=FALSE, message=FALSE, fig.height=3.5, fig.width=6}
set.seed(20)

#simulate the predictor variable as before
x <- rnorm(100)

#calculate the log mean of the model
log.mu <- 0.5 + 0.2*x

# calculate the response variables (y) from the possion distn.
y <- rpois(100, exp(log.mu))

```


## Simulate a Generalized Linear Model with Poisson

\scriptsize
```{r,  echo=FALSE, warning=FALSE, message=FALSE, fig.height=3.5, fig.width=6}
dat <- as.tibble(cbind(x,y))
names(dat)[1] <- 'x'
names(dat)[2] <- 'y'


ggplot(dat)+
  geom_point(aes(x=x,y=y), size = 2)

```

