---
title: "Problem Set 10"
author: Due May 6, 2022
output: 
  html_document:
    smart: no
    theme: flatly
    float: yes
    css: "style.css"    
editor_options: 
  chunk_output_type: console
mainfont: Arial
mathfont: Arial
fontsize: 12pt
subtitle: Lauren Sullivan, BIO_SC 8002_01
---


# {.tabset .tabset-pills .tabset-fade}

## Liver Toxicity

```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}

###############
#
# 10 PROBLEM SET
# 
# DUE:        May 6, 2020
# 
# CREATED BY: Lauren Sullivan
# FILES:      "liver.toxicity.csv"
#             "dune.csv"
#             "dune.env.csv"
#
###############


########### setup
rm(list=ls())


library(tidyverse)
library(vegan)  #for standardization, distance matrices and nmds
library(stats)  # for PCA
library(ggbiplot) # for graphing PCA

########### data
liver <- read_csv("../data/liver.toxicity.csv")
dune <- read_csv("../data/dune.csv")
dune.env <- read_csv("../data/dune.env.csv")


```

### Background

For the first part of this homework, we will be using a dataset examining the gene expression that occurs under liver toxicity when rats are placed on four different doses of acetaminophen (50, 150, 1500, and 2000 mg/kg). There are 64 rats in this study, with 16 in each dosage treatment. The datafile is "liver.toxicity.csv". We will look at how the **Y** of gene expression data is influenced by the dosage treatment.  This data comes from a larger R package [looking at gene expression and PCA](https://mixomicsteam.github.io/Bookdown/pca.html#biological-question).  You can find it within `library(mixOmics)`, which is now stored in `library("BiocManager")` - [download it here](https://www.bioconductor.org/packages/release/bioc/html/mixOmics.html).   


**Q1** Let's start by looking at the data.  You will notice that this dataset is extremely wide. So, let's look at the first 20 rows and the first 10 columns of the data. Describe what the data look like (what is in each row, column and cell).

```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}


```

### PCA


**Q2** Run a PCA on the data using `prcomp()`, and summarize the data using `summary()`.  Describe what this tells you. What does the standard deviation row tell you?  How much variance does the 1st, 2nd, and 6th principal component, or axis, describe?


```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}



```


**Q3** Plot this PCA so we can get a better idea of what is going on. Color each point by its dosage group to help visualize patterns. Does it appear that the dosage of acetomenaphin fed to a rat alters its gene expression overall?  I would suggest using `var.axes = FALSE` to remove the arrows on the plot that represent each column because there are too many columns and the figure will be very cluttered. 
```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}


```


**Q4** Let's look at the loading scores for the first 10 genes and the first 4 PC axes. When considering gene A_43_P2128, which PC axis does it load most strongly on (of the first 4 axes)?  What is the direction of this loading? 
```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}



```


**Q5** If you wanted to look at the actual PC values for each axis on each rat, you need to pull out the x values from your pca object (the vale of the rotated data - nonstandardized). Do this for the first 10 rows and the first 4 columns. Describe where rat 3 falls on the PC axes.
```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}


```

**Q6** Now try plotting the PCA data and group the points by the Treatment Groups.  What hypothesis would you draw from the relationships that stand out from this figure?
```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}


```

## Sand Dune Data

### Background

Now we will switch gears a bit and look at a datset of plant species cover (aerial cover of each species within each plot) on dune environments. We have cover data from 20 plots. We also have information on how each site was managed, used (hayfield, haypasture, pasture), it's soil moisture, and it's manuer application levels. Our plant species cover can be found in "dune.csv", while our environmental variables can be found in "dune.env.csv".


### NMDS


**Q7** Run an nMDS model on this dune data, with k=2, and Bray Curtis distance.  Select a standardization and justify your results. Plot your nMDS results, and use hulls to define the different Use types of each plant community.


```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}



```


### Hypothesis Testing


**Q8** With the same Bray Curtis distance, run a permutational ANOVA with `adonis()` to find out if soil moisture ("Moisture") and manure application ("Manure") is a predictor of species composition within these plots.  What do you find?


```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}


```


**Q9** What happens if you keep moisture and manure as predictor variables, but add "Use" as a grouping variable?  What do you find now?

```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}


```


**Q10** What happens if you keep moisture and manure as predictor variables, but add "Use" as a grouping variable?  What do you find now?

```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}



```

**Q11** Finally, test the hypothesis that there is a difference between dispersion among Use groups in this plant community data (see figure from Q7 to look at hulls of Use groups).

```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}

```



